---
title: "Data606 Presentation Question 5.5"
author: "Ritesh_Lohiya"
date: "3/24/2018"
output: html_document
---

A 95% confidence interval for a population mean is given as (18.985, 21.015). This confidence
interval is based on a simple random sample of 36 observations. Calculate the sample mean and
standard deviation. Assume that all conditions necessary for inference are satisfied. Use the
t-distribution in any calculations.


            Lower (18.985)                    Upper (21.015)
          -------------------------|-------------------------

Solution:

UpperBound of Confidence Interval = UB = 21.015

LowerBound of Confidence Interval = LB = 18.985

Sample Mean:

Sample_Mean <- (UB + LB)/2

```{r}
UB <- 21.015
LB <- 18.985
Sample_Mean <- (UB + LB)/2
Sample_Mean
```

Margin of Error:

(UB - LB)/2

```{r}
Margin_of_Error <- (UB - LB)/2
Margin_of_Error
```

Degree of Freedom and T-score:

```{r}
Df <- 36 - 1
T35 <- qt(.95,Df)
T35
```

Standard Deviation:

```{r}
Sd <- ((Margin_of_Error/T35)*sqrt(36))
Sd
```

Validate Results:


```{r}
UB_Val <- 20 + 1.689572 * 3.604462 / sqrt(36)
UB_Val
```

```{r}
LB_Val <- 20 - 1.689572 * 3.604462 / sqrt(36)
LB_Val
```

UB = UB_Val

LB = LB_Val